### 2023-07-05 tesis_Ashish/Chapter2/chapter2.tex
AS thesis update

Luminosity measurement is a crucial aspect of particle physics experiments such as the Compact Muon Solenoid (CMS) experiment at the Large Hadron Collider (LHC). Luminosity refers to the rate at which particles collide within the detector, and accurate measurement of luminosity is essential for precise determination of particle collision rates and cross-sections. The CMS experiment uses several different devices to measure luminosity, including the Pixel Luminosity Telescope (PLT), Fast Beam Condition Monitor (BCM1F), Hadron forward (HF), Drift tubes (DT), radiation monitoring system for environment and safety (RAMSES) as shown in Fig.~\ref{fig:lumino_cms}. The DT luminometer covers the pseudorapidity range from 0 to 1.2, which is the central region of the CMS detector. The PCC luminometer covers the pseudorapidity range from 0 to 2.5, which covers the central and forward regions of the CMS detector. The HF luminometer covers the pseudorapidity range from 3 to 5, which is the most forward region of the CMS detector. The BCM1F luminometer covers the pseudorapidity range from 3.1 to 4.7, which is also in the forward region of the detector. The RAMSES luminometer covers the pseudorapidity range from 4 to 5, which is also in the forward region of the detector. The PLT luminometer covers the pseudorapidity range from 2.8 to 4.7, which is in the forward region of the CMS detector as shown in Fig.~\ref{fig:lumino_cms_pseudo}. In addition, the Pixel Cluster Counting (PCC) and zero counting method are a commonly used techniques for luminosity measurement at the CMS experiment.

@@

\item Pixel Luminosity Telescope (PLT) : The PLT consists of four layers of silicon pixel sensors arranged in a stack, with each layer containing over 8 million individual pixels. The first three layers are located very close to the interaction point where the proton beams collide, and the fourth layer is located further away. The innermost layer is located 2.9 cm from the interaction point. The second layer is located 3.9 cm from the interaction point. The third layer is located 4.9 cm from the interaction point. The outermost layer is located 10.2 cm from the interaction point. The pixel sensors are designed to detect charged particles produced in the collisions as they pass through the detector layers. The PLT operates in what is called the "fast-OR" mode, which detects charged particles produced in a single bunch crossing of the proton beams. In this mode, the PLT triggers on any pixel in the detector that registers a signal above a certain threshold. This signal is called the "fast-OR signal" and indicates that a charged particle has passed through the detector. The PLT uses the information from all four detector planes to reconstruct the trajectories of the charged particles produced in the collisions. This is done by analyzing the signals from the individual pixels and using algorithms to determine the path of the particle through the detector. To measure the luminosity of the collisions, the PLT counts the number of charged particles produced in each bunch crossing, and uses this information to calculate the rate of collisions. The luminosity is then calculated by dividing the collision rate by the cross-section of the proton-proton interactions. The PLT uses a specialized algorithm called the "triple coincidence filter" to identify charged particles that are produced in the collisions and to reject background noise. The filter requires a charged particle to be detected in at least three of the four detector planes, which corresponds to a "triple coincidence" detection. This helps to reduce the contribution of background noise and improve the precision of the luminosity measurement.

\item Fast Beam Conditions Monitor (BCM1F) : The BCM1F is a device used in the CMS (Compact Muon Solenoid) experiment at CERN's Large Hadron Collider (LHC) to measure the luminosity of proton-proton collisions. The luminosity is a measure of the rate of collision events per unit time and per unit area of the beam. The BCM1F is specifically designed to detect the ionization produced by charged particles passing through its diamond sensors, and to convert this signal into a measurement of the beam intensity and luminosity. The BCM1F consists of 24 diamond sensors, each with an active area of 1 cm², arranged in six rings around the beam pipe. The diamond sensors are positioned very close to the CMS detector, just 10 cm away from the beam pipe. This allows the sensors to detect the charged particles produced by the proton-proton collisions. When charged particles pass through the diamond sensors, they ionize the carbon atoms in the diamond crystal, producing electrons and holes. The electrons and holes are then collected by metal electrodes embedded in the diamond, producing a voltage signal proportional to the ionization produced by the charged particles. The voltage signal is then amplified by fast electronics and converted into a digital signal, which is sent to the CMS data acquisition system for further analysis. To measure the luminosity, the BCM1F counts the number of charged particles passing through each diamond sensor per unit time. The signal from each sensor is integrated over a short time interval (usually 25 ns), corresponding to the time it takes for a single proton bunch to pass through the sensors. The integrated signal is then converted into a measure of the instantaneous luminosity, which is proportional to the product of the number of charged particles per bunch and the collision frequency. The number of charged particles per bunch is determined by calibrating the BCM1F with data from other detectors in the CMS experiment, such as the silicon pixel detector and the pixel luminosity telescope. These detectors are used to measure the total number of charged particles produced by the proton-proton collisions, which is then used to calibrate the BCM1F.

@@

\item Hadron Forward Calorimeter: The Hadron Forward (HF) Calorimeter is a critical component of the Compact Muon Solenoid (CMS) experiment at the Large Hadron Collider (LHC) at CERN. It is a subdetector that measures the energy and position of particles produced during proton-proton collisions in the LHC. The HF Calorimeter is designed specifically to measure hadronic particles, which are particles made up of quarks and gluons, in the forward region (i.e., close to the beamline) of the CMS detector. The HF Calorimeter is located at the very front and back of the CMS detector, at a distance of about 11.2 meters from the interaction point. This placement allows it to cover a pseudorapidity range of 2.9 to 5.2, which is important for measuring particles produced at small angles relative to the beamline. The HF Calorimeter is a sampling calorimeter, which means it consists of alternating layers of active and passive materials. The active material, in this case, is scintillating quartz fibers, which emit light when they interact with charged particles. The passive material is steel absorber plates, which serve to initiate hadronic showers, causing the incident hadronic particles to interact and create secondary particles. As these secondary particles pass through the scintillating fibers, they produce light, which is collected by photodetectors and converted into an electrical signal proportional to the particle's energy. Two of the methods used for luminosity measurement in the HF Calorimeter are the Zero Counting (ZC) method and the Missing Energy (ME) method.

%The Zero Counting method is a threshold-based approach for estimating the luminosity. In this method, the total number of inelastic collisions, which produce new particles, is counted by measuring the energy deposited in the HF Calorimeter. To separate inelastic collisions from background noise and other processes, a threshold on the energy deposition is applied. During data-taking, the HF Calorimeter measures the energy deposited in each bunch crossing (when two bunches of protons pass through each other in the LHC). When the energy deposition in both HF Calorimeter sides (HF+ and HF-) is below a certain threshold, it is considered a "zero count" event, which means no inelastic collision has occurred. The number of zero count events is subtracted from the total number of bunch crossings, giving the number of inelastic collisions. The luminosity can then be calculated by dividing the number of inelastic collisions by the integrated time and the inelastic cross-section. The ZC method has the advantage of being robust against pile-up (simultaneous multiple proton-proton collisions) since it relies on counting the events where no energy is deposited. However, it may be affected by detector noise and other background processes that can lead to a non-zero energy deposition.
The zero counting method is another method used by particle colliders to measure the luminosity of collisions. In this method, the luminosity is estimated by counting the number of events with zero particles detected in a specific region of the detector. The basic idea behind the zero counting method is that the probability of having zero particles detected in a given region of the detector is proportional to the probability of having no collisions in that region. This probability can be related to the luminosity of the collisions through the following equation:

@@

where L is the luminosity, N is the number of events with zero particles detected in the region of interest, $N_0$ is the total number of events, $\sigma_{vis}$ is the visible cross-section.One advantage is that it is a simple and robust method that does not require sophisticated detectors or calibration procedures. However, it can be affected by background events, such as cosmic rays or beam-gas interactions, which can produce zero particles in the region of interest and increase the uncertainty in the luminosity measurement.
The Missing Energy method is an alternative approach for luminosity measurement in the HF Calorimeter. This method relies on measuring the total transverse energy (Et) deposited in the calorimeter and comparing it with the expected average transverse energy for inelastic collisions. The idea behind the ME method is that the total transverse energy should be conserved in a collision event. Inelastic collisions are expected to produce a certain amount of transverse energy on average, which can be estimated from simulations and theoretical calculations. The missing energy in an event is defined as the difference between the expected average transverse energy and the measured transverse energy in the HF Calorimeter. For each bunch crossing, the missing energy is calculated, and a distribution of missing energy values is obtained. This distribution is then fitted to a functional form, which includes a term for the inelastic collisions and terms for various background processes (e.g., detector noise, beam-gas interactions). The fit allows extracting the number of inelastic collisions, which can then be used to calculate the luminosity by dividing it by the integrated time and the inelastic cross-section. The ME method is sensitive to the energy scale calibration of the detector and the modeling of background processes. However, it provides an independent and complementary measurement to the ZC method, which helps to reduce systematic uncertainties in the luminosity determination.

@@

\item Muon drift tubes: The DT system is part of the CMS muon system, which also includes the Resistive Plate Chambers (RPCs) and the Cathode Strip Chambers (CSCs). Drift Tubes (DTs) in the CMS experiment are designed to accurately measure the trajectory and momentum of muons produced in proton-proton collisions at the LHC. The DT system is located in the barrel region of the CMS detector, covering the pseudorapidity range of $|\eta| < 1.2$. The CMS Drift Tubes consist of rectangular drift cells filled with a gas mixture, typically of argon and carbon dioxide. Each drift cell contains a thin anode wire held at a high voltage relative to the cell walls, which act as the cathode. When a charged particle like a muon passes through the gas-filled cell, it ionizes the gas molecules, producing free electrons and positively charged ions. These free electrons drift towards the anode wire under the influence of the electric field, while the positive ions drift towards the cathode. The drift time of electrons to the anode wire is measured, and by knowing the drift velocity in the gas mixture, the distance between the ionizing particle's trajectory and the anode wire can be determined. Combining the information from multiple drift cells, the trajectory of the muon can be reconstructed with high precision. It is not a primary method for luminosity measurement, as the system is specifically designed for muon detection and momentum measurement. However, one could potentially use the muon rates detected by the DT system for an indirect luminosity measurement. This would involve correlating the rate of muons detected in specific processes, such as those from W and Z boson decays, with the underlying luminosity. By measuring the rate of these processes and comparing them to the known production cross-sections, the luminosity could be inferred. This approach would require careful control of various experimental factors, such as the efficiency and acceptance of the DT system, the precise determination of the production cross-sections, and corrections for pile-up and background processes. It is important to note that this would not be a primary luminosity measurement technique, but it could provide a complementary cross-check to the main methods used in the CMS experiment.

\item Pixel Cluster Counting (PCC): The pixel detector in the CMS experiment is made up of multiple layers of silicon sensors arranged in a cylindrical shape around the beam pipe. When charged particles from the collisions pass through the silicon sensors, they create electric charge that is read out as a signal. The signals from adjacent sensors are combined to form clusters, which represent the position of the particle track in the detector. The total number of clusters detected over a certain period of time is used to estimate the luminosity. The PCC provides a precise measurement of the luminosity and is particularly useful when the collision rate is high.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.7\textwidth]{ashish_thesis/luminometer_cms.png}
\caption{%
   Compact muon solenoid (CMS) detector with all of the luminosity measurement devices showing PLT, BCM1F, HF, DT, PCC and RAMSES. 
}
\label{fig:lumino_cms}
\end{figure}

\begin{figure}[!htp]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/luminometer_cms_pseudo.png}
\caption{%
    Cross section of the CMS detector in the r-z plane showing all Run 2 luminometer PLT, BCM1F, DTs, and HF location. The two RAMSES monitors used as a luminometer in Run 2 are located directly behind HF. 
}
\label{fig:lumino_cms_pseudo}
\end{figure}

@@

The pixel detector is one of the subdetectors of the CMS experiment that is closest to the interaction point where the proton beams collide. It consists of four barrel layers (L1-L4) at radii of 29, 68, 109 and 160 mm and three forward disks (D1-D3) on each end at distances of 291, 396, and 516 mm of pixel sensors from the center of the detector as shown in Fig.~\ref{fig:phaseI_upgrade}. It is built from 1856 silicon sensor modules, 1184 modules in the barrel layers (BPIX) and 672 modules for the forward disks (FPIX). Modules for barrel layers and forward disks are shown in Fig.~\ref{fig:pixelmodule}.  The modules in BPIX L1 are designed to handle high particle densities and radiation levels. Each module in this layer consists of a matrix of 4160 pixels, arranged in 160 rows and 26 columns. The pixel size in BPIX L1 is approximately 100x150 µm², which allows for high-resolution position measurements of charged particles. The modules in BPIX L2-4 layers are designed to withstand lower radiation levels compared to BPIX L1. Each module in BPIX L2-4 consists of a matrix of 66560 pixels, arranged in 260 rows and 256 columns. The pixel size in BPIX L2-4 is also approximately 100x150 µm². The FPIX modules are also designed to handle high particle densities and radiation levels. Each module in the FPIX consists of a matrix of 16128 pixels, arranged in 160 rows and 101 columns. The pixel size in the FPIX is approximately 100x150 µm². It provide high-precision position measurements of the charged particles produced in the collisions. These position measurements are used to reconstruct the trajectories of the particles and to identify the particles themselves. The pixel detector has a high position resolution, with a spatial resolution of 10-20 microns in the transverse direction and 20-30 microns in the longitudinal direction. The pixel detector also has a high time resolution, with a readout time of 25 nanoseconds. The pixel detector is capable of detecting charged particles with a wide range of energies, from a few hundred MeV to several TeV. The pixel detector is able to operate at very high particle fluxes, with a hit rate of up to $100 MHz/cm^2$. To achieve this high rate capability, the pixel detector uses fast readout electronics, which can read out the signals from the pixels in a few microseconds. The pixel detector is also designed to operate at a high radiation level, with a dose of up to $5 \times 10^{14} neutrons/cm^2$ expected over the lifetime of the detector. To mitigate the effects of radiation damage, the pixel detector uses sensors made of high-purity silicon and a cooling system that keeps the detector at a temperature of -20°C. The pixel detector is also used for the measurement of the luminosity at the CMS experiment. The luminosity measurement is based on counting the number of pixel clusters produced in the detector by the proton-proton collisions. The higher the collision rate, the more clusters are produced in the pixel detector. However, there are several challenges to accurately measure the luminosity with the pixel detector. One challenge is the high rate of particles that pass through the detector, which can cause pileup (multiple particles passing through the detector at the same time). Another challenge is the non-linearity of the pixel detector response, which can lead to inaccuracies in the luminosity measurement. Therefore, the motivation of this work is to develop and test more precise techniques for measuring the luminosity with the pixel detector, in order to improve the accuracy of the luminosity measurement and to better understand the effects of detector response on the measurement.

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/phaseI_upgrade_pixel_detector.png}
\caption{%
   Layout of the CMS Phase-1 pixel detector in longitudinal view. 
}
\label{fig:phaseI_upgrade}
\end{figure}

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/silicon_pixel_module.png}
\caption{%
   Pixel detector modules for BPIX L1 (left), BPIX L2-4 (middle), and
the FPIX detector (right). 
}
\label{fig:pixelmodule}
\end{figure}

\begin{table}
  \begin{center}
    \begin{tabular}{ccccc}  
    \textbf{Layer/Disk}   & \textbf{\# modules} \\ \hline
     L1      &    96        \\  
       L2    &     224   \\ 
        L3   &       352 \\ 
         L4   &       512 \\ 
         D1 inner ring & 88     \\ 
         D1 outer ring &   136    \\ 
         D2 inner ring &      88 \\ 
         D2 outer ring &   136   \\ 
          D3 inner ring&     88  \\ 
         D3 outer ring &    136  \\ 
      \end{tabular}
    \caption{Number of pixel detector modules in each layer and disk}
    \label{tab:modpix}
  \end{center}
\end{table}

@@

The PCC method has several advantages over the other luminosity measurement methods used by the CMS experiment. One advantage is its high precision and stability, as it is based on a well-understood physical process and has a small systematic uncertainty. Another advantage is its independence from the beam conditions, as the pixel detector is insensitive to the beam halo and the beam-gas interactions. The PCC method is also robust against pile-up, which is the presence of multiple proton-proton interactions in a single bunch crossing, as it can distinguish between pixel clusters produced by different collisions. The PCC method is based on the pixel detector, which has a high spatial resolution and can accurately measure the position and energy of charged particles. This allows for precise measurements of the luminosity, even when the particle collision rate is high. Advantages of the PCC method are described below,

@@

\section{Previous studies related to PCC luminosity measurement}

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/Run2018A_old.png}
\caption{%

}
\label{fig:pixelmodule}
\end{figure}

>>>>

%Luminosity measurement is a crucial aspect of particle physics experiments such as the Compact Muon Solenoid (CMS) experiment at the Large Hadron Collider (LHC). Luminosity refers to the rate at which particles collide within the detector, and accurate measurement of luminosity is essential for precise determination of particle collision rates and cross-sections. The CMS experiment uses several different devices to measure luminosity, including the Pixel Luminosity Telescope (PLT), Fast Beam Condition Monitor (BCM1F), Hadron forward (HF), Drift tubes (DT), radiation monitoring system for environment and safety (RAMSES) as shown in Fig.~\ref{fig:lumino_cms}. The DT luminometer covers the pseudorapidity range from 0 to 1.2, which is the central region of the CMS detector. The PCC luminometer covers the pseudorapidity range from 0 to 2.5, which covers the central and forward regions of the CMS detector. The HF luminometer covers the pseudorapidity range from 3 to 5, which is the most forward region of the CMS detector. The BCM1F luminometer covers the pseudorapidity range from 3.1 to 4.7, which is also in the forward region of the detector. The RAMSES luminometer covers the pseudorapidity range from 4 to 5, which is also in the forward region of the detector. The PLT luminometer covers the pseudorapidity range from 2.8 to 4.7, which is in the forward region of the CMS detector as shown in Fig.~\ref{fig:lumino_cms_pseudo}. In addition, the Pixel Cluster Counting (PCC) and zero counting method are a commonly used techniques for luminosity measurement at the CMS experiment.

Luminosity measurement is a crucial aspect of particle physics experiments such as the Compact Muon Solenoid (CMS) at the Large Hadron Collider (LHC). Luminosity signifies the rate at which particles collide per unit area per second, making its precision measurement vital for the precise determination of particle collision rates and cross-sections. The CMS experiment employs several devices to measure luminosity, each covering specific pseudorapidity ranges and regions within the CMS detector \cite{CMS-PAS-LUM-18-002}. The use of several luminometers as shown in Fig.~\ref{fig:lumino_cms} offers a multitude of advantages. Primarily, it ensures redundancy and acts as a backup for potential device failures, ensuring that data collection can proceed without interruption. Additionally, having multiple luminometers allows for cross-checking, providing greater consistency in results and allowing for more reliable data interpretation. These different devices enable better control over systematic uncertainties, as any discrepancies can be identified and adjusted for. Moreover, each device employs different method to measure luminosity, providing complementary information that enriches the overall understanding of particle interactions. Some luminometers recording low rates can be calibrated using others, enhancing their accuracy and utility. This holistic approach to luminosity measurement significantly improves the robustness and reliability of the data collected. In the central region, The Muon Drift Tubes (DT) luminometer is used covering a pseudorapidity range from 0 to 1.2. The Pixel Cluster Counting (PCC) luminometer also operates in this region and extends into the forward region covering a range from 0 to 2.5. In the forward region, several devices are utilized. The Pixel Luminosity Telescope (PLT) covers the pseudorapidity range from 2.8 to 4.7. The Fast Beam Condition Monitor (BCM1F) functions in a range from 3.1 to 4.7. The Radiation Monitoring System for Environment and Safety (RAMSES) operates in the pseudorapidity range from 4 to 5. In the most Forward Region, The Hadron forward (HF) luminometer caters to the most forward region of the CMS detector \cite{Sirunyan:2759951}, functioning in the pseudorapidity range from 3 to 5  as shown in Fig.~\ref{fig:lumino_cms_pseudo}.

%In addition to these devices, two common techniques used for luminosity measurement at the CMS experiment are the Pixel Cluster Counting (PCC) and zero counting method. Please refer to Figures 1.1 and 1.2 for further visual information about these devices and their respective coverage within the CMS detector.

@@

\item Pixel Luminosity Telescope (PLT) : The PLT consists of four layers of silicon pixel sensors arranged in a stack, with each layer containing over 8 million individual pixels. The first three layers are located very close to the interaction point where the proton beams collide, and the fourth layer is located further away. The innermost layer is located 2.9 cm from the interaction point. The second layer is located 3.9 cm from the interaction point. The third layer is located 4.9 cm from the interaction point. The outermost layer is located 10.2 cm from the interaction point \cite{Ayala:2861814}. The pixel sensors are designed to detect charged particles produced in the collisions as they pass through the detector layers. The PLT operates in what is called the "fast-OR" mode, which detects charged particles produced in a single bunch crossing of the proton beams. In this mode, the PLT triggers on any pixel in the detector that registers a signal above a certain threshold. This signal is called the "fast-OR signal" and indicates that a charged particle has passed through the detector. The PLT uses the information from all four detector planes to reconstruct the trajectories of the charged particles produced in the collisions. This is done by analyzing the signals from the individual pixels and using algorithms to determine the path of the particle through the detector.
  %To measure the luminosity of the collisions, the PLT counts the number of charged particles produced in each bunch crossing, and uses this information to calculate the rate of collisions. The luminosity is then calculated by dividing the collision rate by the cross-section of the proton-proton interactions.
  The PLT uses a specialized algorithm called the "triple coincidence" to identify charged particles that are produced in the collisions and to reject background noise. It requires a charged particle to be detected in at least three of the four detector planes, which corresponds to a "triple coincidence" detection. The rate of triple coincidences is proportional to the luminosity. This helps to reduce the contribution of background noise and improve the precision of the luminosity measurement \cite{Lujan:2017kvh}.

\item Fast Beam Conditions Monitor (BCM1F) : The BCM1F is specifically designed to detect the ionization produced by charged particles passing through its diamond sensors, and to convert this signal into a measurement of the beam intensity and luminosity. The BCM1F consists of 24 diamond sensors, each with an active area of 1 cm², arranged in six rings around the beam pipe. The diamond sensors are positioned very close to the CMS detector, just 10 cm away from the beam pipe. This allows the sensors to detect the charged particles produced by the proton-proton collisions. When charged particles pass through the diamond sensors, they ionize the carbon atoms in the diamond crystal, producing electrons and holes. The electrons and holes are then collected by metal electrodes embedded in the diamond, producing a voltage signal proportional to the ionization produced by the charged particles. The voltage signal is then amplified by fast electronics and converted into a digital signal, which is sent to the CMS data acquisition system for further analysis. To measure the luminosity, the BCM1F counts the number of charged particles passing through each diamond sensor per unit time. The signal from each sensor is integrated over a short time interval (usually 25 ns), corresponding to the time it takes for a single proton bunch to pass through the sensors. The integrated signal is then converted into a measure of the instantaneous luminosity, which is proportional to the product of the number of charged particles per bunch and the collision frequency. The number of charged particles per bunch is determined by calibrating the BCM1F with data from other detectors in the CMS experiment, such as the silicon pixel detector and the pixel luminosity telescope. These detectors are used to measure the total number of charged particles produced by the proton-proton collisions, which is then used to calibrate the BCM1F \cite{CMS-DP-2013-003}.

@@

\item Hadron Forward Calorimeter: The Hadron Forward (HF) Calorimeter is a critical component of the Compact Muon Solenoid (CMS) experiment at the Large Hadron Collider (LHC) at CERN. It is a subdetector that measures the energy, position and arrival time of particles produced during proton-proton collisions in the LHC. The HF Calorimeter is designed specifically to measure hadron particles, which are particles made up of quarks and gluons, in the forward region (i.e., close to the beamline) of the CMS detector. The HF Calorimeter is located at the very front and back of the CMS detector, at a distance of about 11.2 meters from the interaction point. This placement allows it to cover a pseudorapidity range of 2.9 to 5.2, which is important for measuring particles produced at small angles relative to the beamline. The HF Calorimeter is a sampling calorimeter, which means it consists of alternating layers of active and passive materials. The active material, in this case, is scintillating quartz fibers, which emit light when they interact with charged particles. The passive material is steel absorber plates, which serve to initiate hadron showers, causing the incident hadron to interact and create secondary particles. As these secondary particles pass through the scintillating fibers, they produce light, which is collected by photodetectors and converted into an electrical signal proportional to the particle's energy. Two separate algorithm are used for luminosity measurement in the HF Calorimeter, HFOC and HFET based on the Zero Counting (ZC) method and the Missing Transverse Energy (MET) methods respectively \cite{CMS-PAS-LUM-15-001}.

%The Zero Counting method is a threshold-based approach for estimating the luminosity. In this method, the total number of inelastic collisions, which produce new particles, is counted by measuring the energy deposited in the HF Calorimeter. To separate inelastic collisions from background noise and other processes, a threshold on the energy deposition is applied. During data-taking, the HF Calorimeter measures the energy deposited in each bunch crossing (when two bunches of protons pass through each other in the LHC). When the energy deposition in both HF Calorimeter sides (HF+ and HF-) is below a certain threshold, it is considered a "zero count" event, which means no inelastic collision has occurred. The number of zero count events is subtracted from the total number of bunch crossings, giving the number of inelastic collisions. The luminosity can then be calculated by dividing the number of inelastic collisions by the integrated time and the inelastic cross-section. The ZC method has the advantage of being robust against pile-up (simultaneous multiple proton-proton collisions) since it relies on counting the events where no energy is deposited. However, it may be affected by detector noise and other background processes that can lead to a non-zero energy deposition.
The zero counting method is used by particle colliders to measure the luminosity. In this method, the luminosity is estimated by counting the number of events with zero particles detected in a specific region of the detector. The basic idea behind the zero counting method is that the probability of having zero particles detected in a given region of the detector is proportional to the probability of having no collisions in that region. This probability can be related to the luminosity of the collisions through the following equation:

@@

where L is the luminosity, N is the number of events with zero particles detected in the region of interest, $N_0$ is the total number of events, $\sigma_{vis}$ is the visible cross-section. One advantage is that it is a simple and robust method that does not require sophisticated detectors or calibration procedures. However, it can be affected by background events, such as cosmic rays or beam-gas interactions, which can produce zero particles in the region of interest and increase the uncertainty in the luminosity measurement. \\
The Missing Energy method is an alternative approach for luminosity measurement in the HF Calorimeter. This method relies on measuring the total transverse energy (Et) deposited in the calorimeter and comparing it with the expected average transverse energy for inelastic collisions. The idea behind the ME method is that the total transverse energy should be conserved in a collision event. Inelastic collisions are expected to produce a certain amount of transverse energy on average, which can be estimated from simulations and theoretical calculations. The missing energy in an event is defined as the difference between the expected average transverse energy and the measured transverse energy in the HF Calorimeter. For each bunch crossing, the missing energy is calculated, and a distribution of missing energy values is obtained. This distribution is then fitted to a functional form, which includes a term for the inelastic collisions and terms for various background processes (e.g., detector noise, beam-gas interactions). The fit allows extracting the number of inelastic collisions, which can then be used to calculate the luminosity by dividing it by the integrated time and the inelastic cross-section. The ME method is sensitive to the energy scale calibration of the detector and the modeling of background processes. However, it provides an independent and complementary measurement to the ZC method, which helps to reduce systematic uncertainties in the luminosity determination \cite{Sirunyan:2759951}.

@@

\item Muon drift tubes: The DT system is part of the CMS muon system, which also includes the Resistive Plate Chambers (RPCs) and the Cathode Strip Chambers (CSCs). Drift Tubes (DTs) in the CMS experiment are designed to accurately measure the trajectory and momentum of muons produced in proton-proton collisions at the LHC. The DT system is located in the barrel region of the CMS detector, covering the pseudorapidity range of $|\eta| < 1.2$. The CMS Drift Tubes consist of rectangular drift cells filled with a gas mixture, typically of argon and carbon dioxide. Each drift cell contains a thin anode wire held at a high voltage relative to the cell walls, which act as the cathode. When a charged particle like a muon passes through the gas-filled cell, it ionizes the gas molecules, producing free electrons and positively charged ions. These free electrons drift towards the anode wire under the influence of the electric field, while the positive ions drift towards the cathode. The drift time of electrons to the anode wire is measured, and by knowing the drift velocity in the gas mixture, the distance between the ionizing particle's trajectory and the anode wire can be determined. Combining the information from multiple drift cells, the trajectory of the muon can be reconstructed with high precision. It is not a primary method for luminosity measurement, as the system is specifically designed for muon detection and momentum measurement. However, one could potentially use the muon rates detected by the DT system for an indirect luminosity measurement. This would involve correlating the rate of muons detected in specific processes, such as those from W and Z boson decays, with the underlying luminosity. By measuring the rate of these processes and comparing them to the known production cross-sections, the luminosity could be inferred. This approach would require careful control of various experimental factors, such as the efficiency and acceptance of the DT system, the precise determination of the production cross-sections, and corrections for pile-up and background processes. It is important to note that this would not be a primary luminosity measurement technique, but it could provide a complementary cross-check to the main methods used in the CMS experiment \cite{Chatrchyan:2013sba}.

\item Pixel Cluster Counting (PCC): The pixel detector in the CMS experiment is made up of multiple layers of silicon sensors arranged in a cylindrical shape around the beam pipe. When charged particles from the collisions pass through the silicon sensors, they create electric charge that is read out as a signal. The signals from adjacent sensors are combined to form clusters, which represent the position of the particle track in the detector. The total number of clusters detected over a certain period of time is used to estimate the luminosity. The PCC provides a precise measurement of the luminosity and is particularly useful when the collision rate is high \cite{CMS-PAS-LUM-13-001}.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/luminometer_cms.png}
\caption[CMS luminometers]{%
   Compact muon solenoid (CMS) detector with all of the luminosity measurement devices showing PLT, BCM1F, HF, DT, PCC and RAMSES. 
}
\label{fig:lumino_cms}
\end{figure}

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/luminometer_cms_pseudo.png}
\caption[Location of CMS luminometers]{%
    Cross section of the CMS detector in the r-z plane showing all Run 2 luminometer PLT, BCM1F, DTs, and HF location. The two RAMSES monitors used as a luminometer in Run 2 are located directly behind HF. 
}
\label{fig:lumino_cms_pseudo}
\end{figure}

@@

The pixel detector is one of the subdetectors of the CMS experiment that is closest to the interaction point where the proton beams collide. It consists of four barrel layers (L1-L4) at radii of 29, 68, 109 and 160 mm and three forward disks (D1-D3) on each end at distances of 291, 396, and 516 mm of pixel sensors from the center of the detector as shown in Fig.~\ref{fig:phaseI_upgrade}. It is built from 1856 silicon sensor modules, 1184 modules in the barrel layers (BPIX) and 672 modules for the forward disks (FPIX) \cite{Adam:2748381}. Modules for barrel layers and forward disks are shown in Fig.~\ref{fig:pixelmodule}. Table~\ref{tab:modpix} shows the number of modules in each pixel layer and disk. The modules in BPIX L1 are designed to handle high particle densities and radiation levels. Each module in this layer consists of a matrix of 4160 pixels, arranged in 160 rows and 26 columns. The pixel size in BPIX L1 is approximately 100x150 µm², which allows for high-resolution position measurements of charged particles. The modules in BPIX L2-4 layers are designed to withstand lower radiation levels compared to BPIX L1. Each module in BPIX L2-4 consists of a matrix of 66560 pixels, arranged in 260 rows and 256 columns. The pixel size in BPIX L2-4 is also approximately 100x150 µm². The FPIX modules are also designed to handle high particle densities and radiation levels. Each module in the FPIX consists of a matrix of 16128 pixels, arranged in 160 rows and 101 columns. The pixel size in the FPIX is approximately 100x150 µm². It provide high-precision position measurements of the charged particles produced in the collisions. These position measurements are used to reconstruct the trajectories of the particles and to identify the particles themselves. The pixel detector has a high position resolution, with a spatial resolution of 10-20 microns in the transverse direction and 20-30 microns in the longitudinal direction. The pixel detector also has a high time resolution, with a readout time of 25 nanoseconds. The pixel detector is capable of detecting charged particles with a wide range of energies, from a few hundred MeV to several TeV. The pixel detector is able to operate at very high particle fluxes, with a hit rate of up to $100 MHz/cm^2$. To achieve this high rate capability, the pixel detector uses fast readout electronics, which can read out the signals from the pixels in a few microseconds. The pixel detector is also designed to operate at a high radiation level, with a dose of up to $5 \times 10^{14} neutrons/cm^2$ expected over the lifetime of the detector. To mitigate the effects of radiation damage, the pixel detector uses sensors made of high-purity silicon and a cooling system that keeps the detector at a temperature of -20°C. The pixel detector is also used for the measurement of the luminosity at the CMS experiment. The luminosity measurement is based on counting the number of pixel clusters produced in the detector by the proton-proton collisions. The higher the collision rate, the more clusters are produced in the pixel detector. However, there are several challenges to accurately measure the luminosity with the pixel detector. One challenge is the high rate of particles that pass through the detector, which can cause pileup (multiple particles passing through the detector at the same time). Another challenge is the non-linearity of the pixel detector response, which can lead to inaccuracies in the luminosity measurement. Therefore, the motivation of this research work is to develop and test more precise techniques for measuring the luminosity with the pixel detector, in order to improve the precision of the luminosity measurement and to better understand the effects of detector response on the measurement.

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/phaseI_upgrade_pixel_detector.png}
\caption[Pixel detector layout]{%
   Layout of the CMS Phase-1 upgraded pixel detector in longitudinal view. 
}
\label{fig:phaseI_upgrade}
\end{figure}

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/silicon_pixel_module.png}
\caption[Design of pixel module]{%
   Pixel detector modules for BPIX L1 (left), BPIX L2-4 (middle), and
the FPIX detector (right). 
}
\label{fig:pixelmodule}
\end{figure}

\begin{table}
  \begin{center}
    \begin{tabular}{ccccc}  
    \textbf{Layer/Disk}   & \textbf{\# modules} \\ \hline
     L1      &    96        \\  
       L2    &     224   \\ 
        L3   &       352 \\ 
         L4   &       512 \\ 
         D1 inner ring & 88     \\ 
         D1 outer ring &   136    \\ 
         D2 inner ring &      88 \\ 
         D2 outer ring &   136   \\ 
          D3 inner ring&     88  \\ 
         D3 outer ring &    136  \\ 
      \end{tabular}
    \caption[\# modules in pixel layers and disks]{Number of pixel detector modules in each layer and disk.}
    \label{tab:modpix}
  \end{center}
\end{table}

@@

The PCC method has several advantages over the other luminosity measurement methods used by the CMS experiment. One advantage is its high precision and stability, as it is based on a well-understood physical process and has a small systematic uncertainty. %Another advantage is its independence from the beam conditions, as the pixel detector is insensitive to the beam halo and the beam-gas interactions.
The PCC method is also robust against pile-up, which is the presence of multiple proton-proton interactions in a single bunch crossing, as it can distinguish between pixel clusters produced by different collisions. The PCC method is based on the pixel detector, which has a high spatial resolution and can accurately measure the position and energy of charged particles. This allows for precise measurements of the luminosity, even when the particle collision rate is high. Advantages of the PCC method are described below,

@@

\newpage
\section{Previous studies on PCC luminosity measurement}
